---
title: "Trait Annotation Backbone"
author: "Oceana Tavasieff"
date: "9/10/2021"
output: html_document
---

Your input: -table of community data with abundance and grouped by tree
Your output: -two dataframes with trees as column names, relativized abundances by tree. Columns are either OTU's or trait-complexes. Output as .rds files
*Your notes below*

**Erlandson dataset. 10 Sept 2021. Must re-do with new trait-complexes (to avoid double-counting tips). Instead of manufacturing a single unknown column, we are separating tips based on how tey were able to be annotated. 
For the first part, that's foraging distance (short, medium, long, unknown). 
For the second, that's rhizomorph presence (rhizo, norhizo, unknown)
E.g. medium_rhizo, unknown_norhizo, unknown_unknown

This should preserve the trends we noticed before and certainly make some tables easier to create. **

#SETUP
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

```{r read in FungalTraits doc}
short_traits <- read_xlsx("~/functional-fungi/FungalTraits 1.2_ver_16Dec_2020.xlsx") %>% rename(Genus = GENUS) 

ecm_short <- short_traits %>% drop_na(Ectomycorrhiza_exploration_type_template) %>% arrange(Genus)
#colnames(short_traits)
```

```{r Install Fun^Fun db which may have more consistent annotations}
# Version: https://zenodo.org/record/1216257

# Via: https://github.com/traitecoevo/fungaltraits

#install.packages("devtools")
#devtools::install_github("ropenscilabs/datastorr")
#devtools::install_github("traitecoevo/fungaltraits")
library(fungaltraits)
```

```{r Access Fun^Fun db, include=FALSE}
#fun_fun_db <-fungal_traits()
long_traits <- fungal_traits() %>% separate(speciesMatched, c("Genus", "Species", "Subspecies", "Etc"), " ")

# Add an exploration type column from the fungaltraits database 
exp_type_1 <- long_traits %>% select(Genus, Species, em_expl)
exp_type_2 <- short_traits %>% select(Genus, Ectomycorrhiza_exploration_type_template)


exp_type <- merge(exp_type_1, exp_type_2) %>% unite(EXP.TYPE, em_expl, Ectomycorrhiza_exploration_type_template, sep = "_", na.rm=T) %>% unite(ID, Genus, Species, sep=" ", remove = F)

#fun_fun_traits <- colnames(fungal_traits())
 
```


#YOUR INPUT
-Community data
```{r}
salix_raw <- read.csv("ccectos_wtaxonomy.csv")
salix <- salix_raw %>% select(-X)
erlandson_annotated <- read.csv("salix_rhizomorph.csv") %>%  mutate(Species = ifelse(Species %in% "", "sp.", Species)) %>% unite(Ecto_ID, Genus:Species, sep = " ", remove = FALSE)

#Merge sample data and our manually DEEMY annotated (and BLAST-assigned) 

erl_DEEMY <- left_join(salix, erlandson_annotated) %>% select(Tip:Genus, Plspp, mean, Ecto_ID, Rhizomorph, Foraging_distance)%>% 
  mutate(enviro = if_else(mean < (-.5), "arid", "mesic"),
         site = rep("cedar_creek"),
         ID=Ecto_ID) 
```

```{r Annotate Source Dataset}
erl_tbAnnotated <- erl_DEEMY %>% select(Tip, Genus, Ecto_spp, ID, Plspp, Rhizomorph, Foraging_distance)

#Below is annotated dataframe
erl_ann <- left_join(erl_tbAnnotated, exp_type) %>%  drop_na(Ecto_spp) %>% mutate(
  distance =  case_when(grepl("long", Foraging_distance) ~ "long",
                        grepl("long", EXP.TYPE) ~ "long",
                        grepl("L", EXP.TYPE) ~ "long",
                        grepl("medium", Foraging_distance)~"medium",
                        grepl("medium", EXP.TYPE) ~ "medium",
                        grepl("M", EXP.TYPE) ~ "medium",
                        grepl("short", Foraging_distance) ~ "short",
                        grepl("short", EXP.TYPE) ~ "short",
                        grepl("S", EXP.TYPE) ~ "short",
                        grepl("contact", Foraging_distance)~"short",
                        grepl("contact", EXP.TYPE) ~ "short",
                        grepl("C", EXP.TYPE)~"short"),
  NoTips=rep(1)) %>% distinct(Tip, .keep_all = T) %>% select(Ecto_spp, Plspp, distance, Rhizomorph, NoTips) 
```

```{r Generate Trait Matrix}

erl_un <- erl_ann
erl_un$distance <- replace_na(erl_un$distance, "unknown")
erl_un$Rhizomorph <- replace_na(erl_un$Rhizomorph, "unknown")
erl_trait <- erl_un %>% mutate(
  Rhizomorph = ifelse(Rhizomorph == "Y", "rhizo", Rhizomorph),
  Rhizomorph = ifelse(Rhizomorph == "N", "norhizo", Rhizomorph),
  Rhizomorph = ifelse(Rhizomorph == "", "unknown", Rhizomorph))%>% unite(Trait_species, distance, Rhizomorph, sep = "_" ) %>% mutate(Trait_species = as.character(.$Trait_species)) %>% group_by(Plspp, Trait_species) %>% summarise(count=n()) %>% pivot_wider(., names_from = Trait_species, values_from = count) %>%  mutate_all(., ~replace(., is.na(.), 0)) %>% column_to_rownames(var="Plspp")

#RELATAVIZE

erl_rel_trait <- erl_trait %>% mutate(sumAbund = select(., c(1:7,)) %>% 
             rowSums(na.rm = TRUE)) %>% 
  relocate(sumAbund, .before = medium_rhizo)

for(i in 2:ncol(erl_rel_trait)) {
  erl_rel_trait[,i] <- erl_rel_trait[,i]/erl_rel_trait$sumAbund
}

erl_rel_trait <- erl_rel_trait %>% mutate(sumAbund = NULL)

saveRDS(erl_rel_trait, "NMDS_4_input/erlandson_trait_matrix.rds")
```


```{r Generate OTU Table}
erl_OTU <- erl_ann %>% select(Ecto_spp, Plspp, NoTips) %>% group_by(Plspp, Ecto_spp) %>% summarize(count=n()) %>% pivot_wider(., names_from = Ecto_spp, values_from = count) %>% mutate_all(., ~replace(., is.na(.), 0)) %>% column_to_rownames(var="Plspp")

#RELATAVIZE

erl_rel_OTU <- erl_OTU %>% mutate(sumAbund = select(., c(1:81,)) %>% 
             rowSums(na.rm = TRUE)) %>% 
  relocate(sumAbund, .before = Otidea_cf._bufonia)

for(i in 2:ncol(erl_rel_OTU)) {
  erl_rel_OTU[,i] <- erl_rel_OTU[,i]/erl_rel_OTU$sumAbund
}

erl_rel_OTU <- erl_rel_OTU %>% mutate(sumAbund = NULL)

saveRDS(erl_rel_OTU, "NMDS_4_input/erlandson_OTU_matrix.rds")
```

```{r set aside tables for LB analysis}

erl_list_enviro <- erl_DEEMY %>% select(Plspp, mean, enviro, site) %>% distinct(.)
saveRDS(erl_list_enviro, "Tables_for_LB_Pairwise/erlandson_enviro_list.rds")
saveRDS(erl_trait, "Tables_for_LB_Pairwise/erlandson_traits.rds")
saveRDS(erl_OTU,"Tables_for_LB_Pairwise/erlandson_OTUs.rds" )
```

