---
title: "Erlandson Abundance Table Generation"
author: "Oceana"
date: "4/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
```


```{r}
#Read in OTU data (better our annotated taxa list, not the metadata?)

erlandson_annotated <- read.csv("salix_rhizomorph.csv") %>%  mutate(Species = ifelse(Species %in% "", "sp.", Species)) %>% unite(Ecto_ID, Genus:Species, sep = " ", remove = FALSE)
erlandson_taxa = read_csv("ccectos_wtaxonomy.csv")

erl_tax_an <- left_join(erlandson_taxa, erlandson_annotated) %>% select(Plspp, Ecto_ID, Ecto_spp) %>%  mutate(No.Tip = rep(1)) %>% group_by(Plspp, Ecto_spp) %>% summarize (No.Tips = n())

OTU <- erl_tax_an %>% select(Plspp,Ecto_spp, No.Tips) %>%pivot_wider(names_from = Ecto_spp, values_from = No.Tips) %>% column_to_rownames(var = "Plspp")
```

```{r}

saveRDS(OTU, file = "erl_OTU.rds")


#erlandson_taxa = read_csv("ccectos_wtaxonomy.csv") %>% select(Ecto_spp) %>% distinct() #check that we have same # of OTU's
```

Generate Abudncance table from erlandson raw dataset

ID by tree ("Ind" column)
Relativize ectoID by tree (eg communuity at tree/site was 20% (.20 or 20?) Laccaria) and by counts of Ecto_spp and divide by number of unique root tips at the tree (~12/tree, expect fewer)

spp matrix = per tree abund data
trait matrix = per tree trait data

-paste to combo.matrix your data to generate species NMDS 
Structure of end dataset is:
  - 
  
# Laura's Code
# Trying out a new pipeline for the Erlandson data.


```{r include=FALSE}
# Trying out a new pipeline for the Erlandson data.

library(tidyverse)

# Goal: Get 1) a table of relative OTU abundances by tree,
# 2) a table of relative TRAIT abundances by tree.
# Push both of those through an NMDS
# Color code by site aridity

#setwd("/Users/laurabogar/Documents/2020-2021/Oceana research/TraitAnalysisforOceana")

erlandson_taxa = read_csv("ccectos_wtaxonomy.csv")

# Relative abundance by tree:

# Several root tips don't have tree-level identifiers, remove.
erlandson_taxa = erlandson_taxa %>% filter(!is.na(Ind))

# Group by tree individual?
totalabund = erlandson_taxa %>% group_by(Ind, Ecto_spp) %>% summarize(count = n())

relabund = totalabund %>% group_by(Ind) %>% summarize(totaltips = sum(count))
# No, there are a bunch of trees that only have a couple of root 

# Let's group by tree species and site (Plspp.)
totalabund = erlandson_taxa %>% group_by(Plspp, Ecto_spp) %>% summarize(count = n())
bysiteandtree = totalabund %>% group_by(Plspp) %>% summarize(totaltips = sum(count))

together = left_join(totalabund, bysiteandtree)

relabund = together %>% mutate(relabund = count/totaltips)

justrelabund = relabund %>% select(Tree_site = Plspp, Ecto_spp, relabund)


erlandson_otus = justrelabund %>% spread(Ecto_spp, relabund)
erlandson_otus[is.na(erlandson_otus)] = 0  # I'm sure there's a tidyr way to do this but I don't know it. ## take anyhing that was printed as NA as a zero (bc we should be reading these as zeros)

erlandson_otus_for_NMDS = as.matrix(erlandson_otus)


# Note: You will probably have to remove the Tree_site column before
# merging with other data (Tejon, Mendocino) for metaMDS().

# Now to repeat for the trait data!

# Note: I am not sure how Holly relativized the Tejon and Mendocino data
# so that they were comparable to each other. They weren't relative abundance, whatever
# was happening. I will ask. We should use whatever metric she used while pulling other
# data into the analysis.
```

# My code, Generate Trait Table abund Matrix
```{r include=FALSE}
trait_input <- read.csv("erlandson_matrix_input.csv") %>% filter(!is.na(type)) %>% group_by(Plspp, type)%>% summarize(count=n())

trait_relabund_join <- left_join(relabund, trait_input) %>% filter(!is.na(type)) 

filter_trait <- trait_relabund_join %>%  group_by(Tree_site=Plspp, type) %>% select(Tree_site, type, Ecto_spp, relabund) %>% summarize(mysum = sum(relabund))



#ADD identical traits/site together (i.e. if there are multiple "mediums" in the same Tree_site, combine relabund and type rows)



trait_otu <- filter_trait %>% spread(type, mysum)

trait_otu[is.na(trait_otu)] = 0 
trait_otu_NMDS <- as.matrix(trait_otu)


```

```{r include=FALSE}
write.csv(erlandson_otus_for_NMDS, "erlandson_OTU.csv")
write.csv(trait_otu_NMDS, "trait_OTU.csv")
```

