---
title: "Trait Annotation Backbone"
author: "Oceana Tavasieff"
date: "9/10/2021"
output: html_document
---

Your input: -table of community data with abundance and grouped by tree
Your output: -two dataframes with trees as column names, relativized abundances by tree. Columns are either OTU's or trait-complexes. Output as .rds files
*Your notes below*
**Mendocino dataset**
#SETUP
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```
```{r read in FungalTraits doc}
short_traits <- read_xlsx("~/functional-fungi/FungalTraits 1.2_ver_16Dec_2020.xlsx") %>% rename(Genus = GENUS) 

ecm_short <- short_traits %>% drop_na(Ectomycorrhiza_exploration_type_template) %>% arrange(Genus)
#colnames(short_traits)
```

```{r Install Fun^Fun db which may have more consistent annotations}
# Version: https://zenodo.org/record/1216257

# Via: https://github.com/traitecoevo/fungaltraits

#install.packages("devtools")
#devtools::install_github("ropenscilabs/datastorr")
#devtools::install_github("traitecoevo/fungaltraits")
library(fungaltraits)
```

```{r Access Fun^Fun db, include=FALSE}
#fun_fun_db <-fungal_traits()
long_traits <- fungal_traits() %>% separate(speciesMatched, c("Genus", "Species", "Subspecies", "Etc"), " ")

# Add an exploration type column from the fungaltraits database 
exp_type_1 <- long_traits %>% select(Genus, Species, em_expl)
exp_type_2 <- short_traits %>% select(Genus, Ectomycorrhiza_exploration_type_template)


exp_type <- merge(exp_type_1, exp_type_2) %>% unite(EXP.TYPE, em_expl, Ectomycorrhiza_exploration_type_template, sep = "_", na.rm=T) 

fun_fun_traits <- colnames(fungal_traits())
 
```


#YOUR INPUT
```{r}
mendo_metadata <- read.csv("TraitAnalysisforOceana/Curated_Assembly_3July2012.csv")
```
## Mendocino Data Cleaning (Holly)
```{r}
#read in files
pygmy.raw<-read.csv("TraitAnalysisforOceana/Curated_Assembly_3July2012.csv")
#head(pygmy.raw) #looks at first 6 lines
#names(pygmy.raw) #looks at variable names
pygmy.clean<-read.csv("TraitAnalysisforOceana/pygmy-cleandata-wsoils.csv")
#is.na(tail(pygmy.clean))
pygmy.crop <- pygmy.clean %>% group_by(SpeciesCode) %>% 
  arrange(., SpeciesCode) %>% ungroup()
```
```{r}
#############Condensing to core by species
pygmy.core<-aggregate(pygmy.clean$No.Tips,
                      list(pygmy.clean$Sample.ID,
                           pygmy.clean$SpeciesCode,
                           pygmy.clean$Tree,
                           pygmy.clean$Core,
                           pygmy.clean$Terrace,
                           pygmy.clean$Pygmy,
                           pygmy.clean$Trait.ForagingType,
                           pygmy.clean$Trait.Hydrophobicity,
                           pygmy.clean$Trait.Rhizomorphs,
                           pygmy.clean$Trait.HostTree)
                      ,sum)

colnames(pygmy.core)<-c("Sample.ID","SpeciesCode","Tree","Core","Terrace","Pygmy","Trait.ForagingType","Trait.Hydrophobicity","Trait.Rhizomorphs","Trait.HostTree","Tips")
# ^ the above works better than what I can do
#head(pygmy.core)

pygmy.core<-cbind(pygmy.core,c(rep(1,dim(pygmy.core)[1])))

colnames(pygmy.core)<-c("Sample.ID","SpeciesCode","Tree","Core","Terrace","Pygmy","Trait.ForagingType","Trait.Hydrophobicity","Trait.Rhizomorphs","Trait.HostTree","Tips","Spp")


#############Condensing to tree by species
names(pygmy.clean)

pygmy.tree<-aggregate(pygmy.clean$No.Tips,
                  list(pygmy.clean$SpeciesCode,
                    pygmy.clean$Tree,
                    pygmy.clean$Terrace,
                    pygmy.clean$Pygmy,
                    pygmy.clean$Trait.ForagingType,
                    pygmy.clean$Trait.Hydrophobicity,
                    pygmy.clean$Trait.Rhizomorphs,
                    pygmy.clean$Trait.HostTree)
                  ,sum)

#colnames(pygmy.core)<-c("SpeciesCode","Tree","Terrace","Pygmy","Trait.ForagingType","Trait.Hydrophobicity","Trait.Rhizomorphs",Tips")

#head(pygmy.tree)

pygmy.tree<-cbind(pygmy.tree,c(rep(1,dim(pygmy.tree)[1])))
colnames(pygmy.tree)<-c("SpeciesCode","Tree","Terrace","Pygmy","Trait.ForagingType","Trait.Hydrophobicity","Trait.Rhizomorphs","Trait.HostTree","Tips","Spp")
tail(pygmy.tree)
```


#Build MENDO tables
```{r}
#Ideally this will happen without Holly's speciesMatrix fxn

#Make Mendo OTU matrix
mendo_OTU_table <- pygmy.tree %>% 
  select(SpeciesCode, Tips, Tree) %>% 
  group_by(Tree) %>% 
  arrange(.,Tree) %>% 
  ungroup(Tree) %>% 
  pivot_wider(., names_from = SpeciesCode, values_from = Tips) %>% 
  replace(., is.na(.), 0)
  
#mendo_OTU_matrix<- as.matrix(mendo_OTU_table)

## Make Mendo TRAIT matrix
mendo_trait_table_foraging <- pygmy.core %>% 
  select(Tree, Trait.ForagingType, Tips) %>% 
  group_by(Tree) %>% 
  arrange(Tree) %>% 
  ungroup() %>% 
  filter(., Trait.ForagingType != "") %>% 
  mutate(row = row_number()) %>%
  pivot_wider(., names_from = Trait.ForagingType, values_from = Tips, values_fill = 0) %>% 
  transmute(Tree = Tree,
            short = shortdistance+contact, 
            medium = mediumdistancefringe+mediumdistancesmooth+mediumdistancemat,
            long = longdistance)

#not 100% sure what row_number does to help pivot.wider--but it works. Thanks stack overflow!

mendo_trait_table_rhizo <- pygmy.core %>% 
  select(Tree, Trait.Rhizomorphs, Tips) %>% 
  group_by(Tree) %>% 
  arrange(Tree) %>% 
  ungroup() %>% 
  filter(., Trait.Rhizomorphs != "") %>% 
    mutate(row = row_number()) %>%
  pivot_wider(., names_from = Trait.Rhizomorphs, values_from = Tips, values_fill = 0) %>% 
  transmute(Tree = Tree,
            norhizo = n,
            rhizo = y)# dropped m column, may have to revisit?

#Investigate "m" values for rhizomorphs
#pygmy_M = filter(pygmy.clean, Trait.Rhizomorphs == "m") # decided these were unknown, and to drop them.

mendo_trait_table <- left_join(mendo_trait_table_foraging, mendo_trait_table_rhizo) %>% 
   group_by(Tree) %>% 
  summarise(short = sum(short),
           medium = sum(medium),
           long = sum(long),
           rhizo = sum(rhizo),
           norhizo = sum(norhizo)) %>% 
     column_to_rownames(., var = "Tree")
#mendo_trait_matrix <- as.matrix(mendo_trait_table)
```


#Relativize MENDO tables
```{r}
#MENDO TRAITS

mendo_rel_traits <- mendo_trait_table %>%
    mutate(sumAbund = select(., short:rhizo) %>% 
             rowSums(na.rm = TRUE)) %>% 
  transmute(short = short/sumAbund, 
         medium = medium/sumAbund,
         long = long/sumAbund,
         norhizo = norhizo/ sumAbund, 
         rhizo = rhizo/sumAbund) 


#MENDO OTU

mendo_rel_OTU <- mendo_OTU_table %>%
    mutate(sumAbund = select(., c(2:83,)) %>% 
             rowSums(na.rm = TRUE)) #%>% 
  #relocate(sumAbund, .after = enviro)

for(i in 2:ncol(mendo_rel_OTU)) {
 mendo_rel_OTU[,i] <- mendo_rel_OTU[,i]/mendo_rel_OTU$sumAbund
  }

mendo_rel_OTU <- mendo_rel_OTU %>% 
  mutate(sumAbund = NULL)

mendo_rel_OTU <- column_to_rownames(mendo_rel_OTU, var = "Tree")  
```

#Generate Mendo Matrices
```{r}
mendo_OTU_matrix <- as.matrix(mendo_rel_OTU)
mendo_trait_matrix <- as.matrix(mendo_rel_traits)

```



-Community data
```{r}
richard_raw <- Richard_et_al_OTU_table <- read_delim("Richard et al OTU table.csv", ";", escape_double = FALSE, trim_ws = TRUE)

```

```{r Generate OTU table}
rich_OTU <- richard_raw %>% select(Taxon, EA07:CS09) %>% mutate(Taxon = str_replace(Taxon, " ", "_")) %>% 
  column_to_rownames(var="Taxon") %>% 
  t %>% 
  as.data.frame %>% 
  rownames_to_column(var ="Taxon") %>% mutate(enviro = if_else(str_detect(Taxon, "E"), "arid", "mesic")) %>% relocate(enviro, .after = Taxon) 

rich_rel_OTU <- rich_OTU %>% mutate(sumAbund = select(., c(3:125,)) %>% 
             rowSums(na.rm = TRUE)) %>% 
  relocate(sumAbund, .after = enviro)

for(i in 4:ncol(rich_rel_OTU)) {
  rich_rel_OTU[,i] <- rich_rel_OTU[,i]/rich_rel_OTU$sumAbund
  }
rich_rel_OTU <- rich_rel_OTU %>% 
  mutate(sumAbund = select(., c(4:126,)) %>% 
             rowSums(na.rm = TRUE)) %>% 
  mutate(sumAbund = NULL,
         enviro = NULL)
rich_rel_OTU <- column_to_rownames(rich_rel_OTU, var = "Taxon")

saveRDS(rich_rel_OTU, "NMDS_4_input/richard_OTU_matrix.rds")

```

```{r Generate Trait Matrix}
rich_4tr <- richard_raw %>% select(-"GenBank accession number", -"Closest GenBank species", -"BLAST expected value‡") %>% separate(Taxon, c("Genus", "Species", "Etc"), " ")

richard_annotated <- left_join(rich_4tr, exp_type) %>% rename(Rhizomorph = "Rhizomorph-DEEMY") %>% unite(ID, Genus, Species, sep=" ") %>% 
  mutate(distance =  case_when(grepl("long", "Foraging_distance-DEEMY") ~ "long",
                               grepl("L", EXP.TYPE) ~ "long",grepl("medium", "Foraging_distance-DEEMY")~"medium",
                               grepl("medium", EXP.TYPE) ~ "medium",
                               grepl("M", EXP.TYPE) ~ "medium",
                               grepl("short", "Foraging_distance-DEEMY") ~ "short",
                               grepl("short", EXP.TYPE) ~ "short",
                               grepl("S", EXP.TYPE) ~ "short",
                               grepl("contact", EXP.TYPE) ~ "short",
                               grepl("C", EXP.TYPE)~"short"))%>% unique() #%>% select(-Etc, -"Foraging_distance-DEEMY", -EXP.TYPE, -Rhizomorph, distance) 
richard_annotated$distance <- replace_na(richard_annotated$distance, "unknown")
richard_annotated$Rhizomorph <- replace_na(richard_annotated$Rhizomorph, "unknown")

tr1 <- richard_annotated  %>% select(ID, Rhizomorph:CS09, distance) %>% group_by(distance, Rhizomorph) %>% summarise_at(vars(EA07:CS09), list(sum)) %>% ungroup() %>% 
  mutate(Rhizomorph = ifelse(Rhizomorph == "y", "rhizo", Rhizomorph),
         Rhizomorph = ifelse(Rhizomorph == "n", "norhizo", Rhizomorph)) %>% unite(Trait_species, distance, Rhizomorph, sep = "_" ) %>% mutate(Trait_species = as.character(.$Trait_species)) %>%  column_to_rownames("Trait_species") 

rich_rel_trait <- tr1 %>%   mutate(sumAbund = rowSums(across(where(is.numeric)))) %>%  mutate(across(everything()), . / sumAbund) %>% select(-sumAbund)


saveRDS(rich_rel_trait, "NMDS_4_input/richard_trait_matrix.rds")
```

