---
title: "Trait Annotation Backbone"
author: "Oceana Tavasieff"
date: "9/10/2021"
output: html_document
---

Your input: -table of community data with abundance and grouped by tree
Your output: -two dataframes with trees as column names, relativized abundances by tree. Columns are either OTU's or trait-complexes. Output as .rds files
*Your notes below*
**Mendocino dataset. Holly's paper (https://academic.oup.com/femsec/article/87/3/797/578450?login=true). we/she used a binomial distinction of if a tree was a pygmy or not (vertical distribution) to determine arid/mesic. **
#SETUP
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```
```{r read in FungalTraits doc, eval=FALSE}
short_traits <- read_xlsx("~/functional-fungi/FungalTraits 1.2_ver_16Dec_2020.xlsx") %>% rename(Genus = GENUS) 

ecm_short <- short_traits %>% drop_na(Ectomycorrhiza_exploration_type_template) %>% arrange(Genus)
#colnames(short_traits)
```

```{r Install Fun^Fun db which may have more consistent annotations, eval=FALSE}
# Version: https://zenodo.org/record/1216257

# Via: https://github.com/traitecoevo/fungaltraits

#install.packages("devtools")
#devtools::install_github("ropenscilabs/datastorr")
#devtools::install_github("traitecoevo/fungaltraits")
library(fungaltraits)
```

```{r Access Fun^Fun db, include=FALSE, eval=FALSE}
#fun_fun_db <-fungal_traits()
long_traits <- fungal_traits() %>% separate(speciesMatched, c("Genus", "Species", "Subspecies", "Etc"), " ")

# Add an exploration type column from the fungaltraits database 
exp_type_1 <- long_traits %>% select(Genus, Species, em_expl)
exp_type_2 <- short_traits %>% select(Genus, Ectomycorrhiza_exploration_type_template)


exp_type <- merge(exp_type_1, exp_type_2) %>% unite(EXP.TYPE, em_expl, Ectomycorrhiza_exploration_type_template, sep = "_", na.rm=T) 

saveRDS(exp_type, "NMDS_4_input/exploration_type_master.rds")

fun_fun_traits <- colnames(fungal_traits())
 
```

```{r read in exploration type dataframe, message=FALSE}
exp_type <- readRDS("exploration_type_master.rds")
```

#YOUR INPUT
```{r}
mendo_metadata <- read.csv("TraitAnalysisforOceana/Curated_Assembly_3July2012.csv")

mendo_tbAnnotated <- mendo_metadata %>% filter(Ectomycorrhizal !="") %>% select(SpeciesCode, Tree, BlastID,Trait.ForagingType, Trait.Rhizomorphs, No.Tips) %>% separate(BlastID, c("Genus", "Species", "Etc"), " ") 

#NM_mendo <- mendo_metadata %>% filter(Ectomycorrhizal != "y")
#Only one missed--gets "m-heliotales" uncertain if this should be included. seems to be ericoid/AM upon googling. 

#Holly includes this one in her analysis

#But Holly does NOT include the auxiliary Foraging type columns. So we won't either.
```

```{r Annotate Source Dataset}

#Below is annotated dataframe
mendo_ann <- left_join(mendo_tbAnnotated, exp_type) %>% mutate(
  distance =  case_when(grepl("long", Trait.ForagingType) ~ "long",
                        grepl("long", EXP.TYPE) ~ "long",
                        grepl("L", EXP.TYPE) ~ "long",
                        grepl("medium", Trait.ForagingType)~"medium",
                        grepl("medium", EXP.TYPE) ~ "medium",
                        grepl("M", EXP.TYPE) ~ "medium",
                        grepl("short", Trait.ForagingType) ~ "short",
                        grepl("short", EXP.TYPE) ~ "short",
                        grepl("S", EXP.TYPE) ~ "short",
                        grepl("contact", Trait.ForagingType)~"short",
                        grepl("contact", EXP.TYPE) ~ "short",
                        grepl("C", EXP.TYPE)~"short")) %>% distinct(SpeciesCode, .keep_all = T) %>% select(Tree, distance, Trait.Rhizomorphs, No.Tips) %>% rename(Rhizomorph=Trait.Rhizomorphs) 
```

```{r Generate Trait Matrix}

mendo_un <- mendo_ann
mendo_un$distance <- replace_na(mendo_un$distance, "unknown")
mendo_un$Rhizomorph <- replace_na(mendo_un$Rhizomorph, "unknown")
mendo_trait <- mendo_un %>% mutate(
  Rhizomorph = ifelse(Rhizomorph == "y", "rhizo", Rhizomorph),
  Rhizomorph = ifelse(Rhizomorph == "n", "norhizo", Rhizomorph),
  Rhizomorph = ifelse(Rhizomorph == "m", "unknown", Rhizomorph),
  Rhizomorph = ifelse(Rhizomorph == "", "unknown", Rhizomorph))%>% unite(Trait_species, distance, Rhizomorph, sep = "_" ) %>% mutate(Trait_species = as.character(.$Trait_species)) %>% group_by(Tree, Trait_species) %>% summarise(count=n()) %>% pivot_wider(., names_from = Trait_species, values_from = count) %>%  mutate_all(., ~replace(., is.na(.), 0)) %>% column_to_rownames(var="Tree")

#RELATAVIZE

mendo_rel_trait <- mendo_trait %>% mutate(sumAbund = select(., c(1:6,)) %>% 
             rowSums(na.rm = TRUE)) %>% 
  relocate(sumAbund, .before = long_rhizo)

for(i in 2:ncol(mendo_rel_trait)) {
  mendo_rel_trait[,i] <- mendo_rel_trait[,i]/mendo_rel_trait$sumAbund
}

mendo_rel_trait <- mendo_rel_trait %>% mutate(sumAbund = NULL)

saveRDS(mendo_rel_trait, "NMDS_4_input/mendo_trait_matrix.rds") 
```

```{r Generate OTU Table}
mendo_OTU <- mendo_tbAnnotated %>% group_by(Tree, SpeciesCode) %>% summarize(count=n())%>% pivot_wider(., names_from = SpeciesCode, values_from = count) %>% mutate_all(., ~replace(., is.na(.), 0)) %>% column_to_rownames(var="Tree")
  
  #RELATAVIZE

mendo_rel_OTU <- mendo_OTU %>% mutate(sumAbund = select(., c(1:82,)) %>% 
             rowSums(na.rm = TRUE)) %>% 
  relocate(sumAbund, .before = "3")

for(i in 2:ncol(mendo_rel_OTU)) {
  mendo_rel_OTU[,i] <- mendo_rel_OTU[,i]/mendo_rel_OTU$sumAbund
}

mendo_rel_OTU <- mendo_rel_OTU %>% mutate(sumAbund = NULL)

saveRDS(mendo_rel_OTU, "NMDS_4_input/mendo_OTU_matrix.rds")
```


```{r set aside tables for LB analysis}
pygmy.clean<-read.csv("TraitAnalysisforOceana/pygmy-cleandata-wsoils.csv")

mendo_list_enviro <- pygmy.clean %>% select(Tree, Pygmy) %>% distinct() %>% mutate(enviro = if_else(Pygmy == 1, "arid", "mesic"),
            Tree = Tree,
            site = rep("mendocino"))
saveRDS(mendo_list_enviro, "Tables_for_LB_Pairwise/mendo_enviro_list.rds")
saveRDS(mendo_trait, "Tables_for_LB_Pairwise/mendo_traits.rds")
saveRDS(mendo_OTU,"Tables_for_LB_Pairwise/mendo_OTUs.rds" )
```